---
title: "LSHW4Paper"
author: "Armand Post"
date: "June 4, 2016"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

---
title: "LiveSessionHW4 Bronx"
author: "Vinh Le"
date: "June 1, 2016"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

## Set working directory and load packages
```{r}
require(gdata)
require(plyr) #Added by Monnie McGee
#install the gdata and plyr packages and load in to R.
library(plyr)
library(gdata)
setwd("/SMU/Doing Data Science/Unit 4/Live Session Files")
```


## Load the datafile
```{r}
bk <- read.csv("rollingsales_bronx.csv",skip=4,header=TRUE)
```


## Clean and format the data
```{r}
bk$SALE.PRICE.N <- as.numeric(gsub("[^[:digit:]]","", bk$SALE.PRICE))
count(is.na(bk$SALE.PRICE.N))
```

```{r}
names(bk) <- tolower(names(bk)) # make all variable names lower case
## Get rid of leading digits
bk$gross.sqft <- as.numeric(gsub("[^[:digit:]]","", bk$gross.square.feet))
bk$land.sqft <- as.numeric(gsub("[^[:digit:]]","", bk$land.square.feet))
bk$year.built <- as.numeric(as.character(bk$year.built))
```


## Below we see the data plotted with $0 sales amounts removed with gross square footage by sales price.
```{r}
## keep only the actual sales

bk.sale <- bk[bk$sale.price.n!=0,]
plot(bk.sale$gross.sqft,bk.sale$sale.price.n,main="Remove $0 sales")
```

## Since the data appeared clustered above, we log transform the data and assess.
```{r}
plot(log10(bk.sale$gross.sqft),log10(bk.sale$sale.price.n),main="Log Transformed, Remove $0 sales")
```



## From the above we can conclude that sales price is positively correlated with gross square footage.


## Next, we Isolate to 1, 2, and 3 family homes.
```{r}
bk.homes <- bk.sale[which(grepl("FAMILY",bk.sale$building.class.category)),]
```


## Below we see the data plotted with 1, 2, and 3 family homes plotted as gross square footage by sales price.
```{r}
plot(bk.homes$gross.sqft,bk.homes$sale.price.n,main="1,2, and 3 family homes, non log transformed")
```


## Since the data appeared clustered above, we log transform the data and assess.
```{r}
plot(log10(bk.homes$gross.sqft),log10(bk.homes$sale.price.n),main="Log Transformed, 1, 2, 3 Family Homes")
summary(bk.homes[which(bk.homes$sale.price.n<100000),])
```


## Since there appear to be a number of outliers, we remove outliers (values above log10 5)
```{r}
## remove outliers that seem like they weren't actual sales
bk.homes$outliers <- (log10(bk.homes$sale.price.n) <=5) + 0
bk.homes <- bk.homes[which(bk.homes$outliers==0),]
plot(log(bk.homes$gross.sqft),log(bk.homes$sale.price.n),main="Non sales removed")
```


## Since the data seems to not have a large positive correlation, we assume that neighborhood location of homes affects prices more than sq ft.
```{r}
boxplot(sale.price.n~neighborhood,data=bk.homes)
```


## However since the number of neighboorhoods is large, data would become thin and difficult to assess.